---
title: "R notes -- Day 1"
author: "Hieu Phay"
date: '2022-05-10'
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Note: We are in a `Rmarkdown` document
- **Bold** _underline_ *italic*

# The philosophy of R

R is a **functional programming language**.

Which means that it operates around the **object-function** metaphor:

- Everything that exists in an R environment is an **object**
- Everything that happens inside an R environment is a **function**

Today we delve into _objects_

## Asssignment
Assignment is done by the sign (`<-`) symbol. 

(`=` works too, but we don't use it).

```{r}
x <- 2
a <- "Thanh"
y <- TRUE
```

Object names must only contain letters and `_`

## Listing and removing objects
Listing objects:

```{r}
ls()
```
Remove objects:

```{r}
skldfjsdklj <- 1
rm(skldfjsdklj)
```


## How to inspect things

You can just type it out!

```{r}
y
```
... or use `str()` (for complicated objects).


# Data types


_Why are objects different? (Thanh)_
Type:
- Boolean: `TRUE` or `FALSE`
- Integer: e.g. `1`, `2`,...
- Float: e.g. `1.0` (is annoying to compares)
- Strings: e.g. `"I am a string!!"`
- ...


Be careful with comparisons using floats!!

```{r}
0.1 + 0.2
```

```{r}
(0.1 + 0.2) == 0.3
```


# Working with vectors

The **atomic** object in R is a `vector` (instead of scalars)

Vector: A list of value in the same **type**.

## Create a vectors

Vectors are created using `c()`

Sequences is created with `:`

```{r}
1:40
```

## Concaternate vectors

`c()` is also concaternate vectors.

```{r}
c(TRUE, 5, 13.5, "Hello")
```
With different types: Data are coerced into the same type.
- Boolean -> Interger -> Float -> String

Almost all functions works with vectors out of the box.

```{r}
c(TRUE, FALSE) + c(3, 4)
```
If vector lengths are uneven

```{r}
1:10 + c(TRUE, FALSE)
```

```{r}
1:10 + c(TRUE, FALSE, FALSE)
```

## Vector indexing

- Index by position number

```{r}
letters[c(1:4, 6:length(letters))]
```
- Index by a logic (boolean) vector

```{r}
letters[letters == "f"]
```

# Puzzle

Part 1: Count the number of times a depth measurement increases from the previous measurement.

```{r}
input <- scan("~/Desktop/R-notes/day-1_puzzle") 
```
Hint: Create a logic vector and use `sum()`

```{r}
sum(c(TRUE, FALSE, TRUE))
```

- Part 2: Count the number of times the 3-period moving sum of depth measurement increases from the previous measurement
