---
title: "R notes -- Day 1"
author: "Hieu Phay"
date: '2022-05-10'
output: 
  pdf_document:
    number_sections: true
    highlight: tango
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R objects

R is a **functional programming language**.

Which means that it operates around the **object-function** metaphor:

-   Everything that exists in an R environment is an **object**
-   Everything that happens inside an R environment is a **function**

Today we delve into *objects*

## Asssignment

Assignment is done by the sign (`<-`) symbol.

(`=` works too, but the convention is that we don't use it for
assignments)

```{r}
x <- 2
a <- "Thanh"
y <- TRUE
```

Object names must only contain letters (uppercase and lowercase),
underscore `_`. Digits is can also be used except at the beginning of
the object name (e.g. `x_1` is fine, but `1_x` is not accepted)

## Listing and removing objects

Listing objects:

```{r}
ls()
```

Remove objects:

```{r}
skldfjsdklj <- 1
rm(skldfjsdklj)
```

## How to inspect things

You can just type it out!

```{r}
y
```

... or use `str()` (for complicated objects -- more on that later).

# Data types

*Question: Why are objects different? (Thanh)*

Some basic data type:

\- Boolean: `TRUE` or `FALSE`

\- Integer: e.g. `1`, `2`,...

\- Float: e.g. `1.0` (is annoying to compares)

\- Strings: e.g. `"I am a string!!"`

\- ...

Be careful with comparisons using floats!! (floating-point errors)

```{r}
0.1 + 0.2
```

```{r}
(0.1 + 0.2) == 0.3
```

# Working with vectors

The **atomic** object in R is a `vector` (instead of scalars)

Vector: A list of value in the same **type**.

## Create a vectors

Vectors are created using `c()`

Sequences is created with `:`

```{r}
1:40
```

## Concaternate vectors

`c()` is also concaternate vectors.

```{r}
c(TRUE, 5, 13.5, "Hello")
```

With different types: Data are coerced into the same type. - Boolean -\>
Interger -\> Float -\> String

Almost all functions works with vectors out of the box.

```{r}
c(TRUE, FALSE) + c(3, 4)
```

If vector lengths are uneven

```{r}
1:10 + c(TRUE, FALSE)
```

```{r}
1:10 + c(TRUE, FALSE, FALSE)
```

## Vector indexing

-   Index by position number

```{r}
letters[c(1:4, 6:length(letters))]
```

-   Index by a logic (boolean) vector

```{r}
letters[letters == "f"]
```

# Puzzle

Puzzle source: [https://adventofcode.com/2021/day/1]()

```{r}
input <- scan("~/Desktop/R-notes/day-1_puzzle") 
```

## Part 1: Count the number of times a depth measurement increases from the previous measurement.

Hint: Create a logic vector and use `sum()`

```{r}
sum(c(TRUE, FALSE, TRUE))
```

**Solution**

Here's the long answer -- using only simple vector solution

```{r}
# Create a lag vector and a lead vector
input_lag <- input[2:length(input)]
input_lead <- input[1:(length(input) - 1)]

# Compare the lag vector with the lead vector
is_increase <- input_lag > input_lead

# Count the number of changes
n_increase <- sum(is_increase)

# Print the result
print(n_increase)
```

In R there are convenient functions for many things. In this case, we
can use the function `diff()`, which compute the absolute differences of
our vector

```{r}
changes <- diff(input)
is_increase <- changes > 0
n_increase <- sum(is_increase)

print(n_increase)
```

## Part 2: Count the number of times the 3-period moving sum of depth measurement increases from the previous measurement

**Solution**

The second part of the puzzle is very similar -- we only have to compute
the 3-period moving sum of our input vector.

```{r}
# Compute the 3-period moving sum
moving_sum3 <- input[1:(length(input) - 2)] +
  input[2:(length(input) - 1)] + input[3:length(input)]


# The following part is similar to Part 1 of the puzzle
changes <- diff(moving_sum3)
is_increase <- changes > 0
n_increase <- sum(is_increase)

print(n_increase)
```
